---
title: "Exercise4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import the raw count table "SRP029880.raw_counts.tsv" from the `compGenomRData` package as a matrix. 
```{r}

```

Apply TPM normalisation to the count matrix.

```{r}

```

Run `colSums(tpm)` to confirm that the sample sizes are all equal to 10^6. 

```{r}

```

Go to https://www.gsea-msigdb.org/gsea/msigdb/collections.jsp#C2 and download the 
'Canonical pathways' gene set collection (2868 gene sets) with gene symbols.

Parse the downloaded gene sets into a named list of vectors 
```{r}

```

Make a histogram of the number of genes per set in the gene set collection
Hint: use ?hist function 
```{r}

```

Calculate the differential gene sets (using `gage` library) between the case samples (columns 1 to 5)
and the control samples (columns 6 to 10) of the normalized (TPM) count matrix.

Hint: see ?gage::gage function. Set compare='as.group' 

```{r}

```

Find the KEGG pathway that shows the most up-regulation in case samples and 
obtain a subset of the tpm normalized count matrix including only the genes from that gene set. 
```{r}

```

Make a heatmap of the most down-regulated gene set
```{r}

```




