---
title: "Exercises-1"
output: html_document
---

The following set of exercises cover:

1. How to download a colorectal cancer dataset from Expression Atlas
2. How to hormalize the count matrix using DESeq2
3. How to do exploratory analysis of the experiment design and count matrix

The diagnostic plots should help you figure out:

"Do the samples cluster by the tissue donor (individual)? 
Alternatively, do they cluster by the type of tissue (normal, tumor, metastatic?)"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Search Expression Atlas database for human colorectal cancer datasets 
```{r}
library(ExpressionAtlas)
datasets <- ExpressionAtlas::searchAtlasExperiments(properties = 'colorectal cancer',
                                                    species = 'human')
```

Download the first dataset in the results (with 54 samples - normal/primary/liver metastases)
Extract count matrix from the downloaded object. 
Extract colData DataFrame from the downloaded object and convert to a data.frame (hint: ?as.data.frame)
```{r}
download <- getAtlasData(experimentAccessions = 'E-GEOD-50760')
exp <- download$`E-GEOD-50760`$rnaseq
counts <- assay(exp, 'counts')
cd <- as.data.frame(colData(exp))
```

Explore the colData data.frame. How many columns are there? Make a summary of each column. 
Hint: see ?table 
```{r}
apply(cd, 2, table)
```

Drop columns that don't have more than one factors (e.g. disease) - are there any other columns with only one factor? Hint: use ?subset function
```{r}
cd <- subset(cd, select = c('AtlasAssayGroup', 'individual', 'organism_part', 'biopsy_site'))
```

Normalize the count matrix using DESeq2
```{r}
dds <- DESeq2::DESeqDataSetFromMatrix(counts, cd, ~ 1)
dds <- estimateSizeFactors(dds)
```

Plot a heatmap of the normalized count matrix using the top most variable 500 genes.
Annotate the columns with available sample annotations in colData object.
```{r}
require(pheatmap)
M <- counts(dds, normalized = T)
plot_heatmap(M, cd)
```

Plot the same heatmap without filtering for top genes (i.e. use all available genes in the normalized 
count matrix). Hint: Don't forget to save your scripts before executing the command!. You'll see why :) 

```{r}

```

Compute the principle components and make PCA plots colored by 
1. AtlasAssayGroup
2. organism_part
3. biopsy_site
Hint: use the plot_pca function defined here: 
https://github.com/BIMSBbioinfo/compgen2020/blob/master/day2_RNAseq/lectureCode/accounting_for_known_covariates_in_DE.R

```{r}
require(ggplot2)
plot_pca(M, cd, by = 'AtlasAssayGroup')
plot_pca(M, cd, by = 'biopsy_site')
plot_pca(M, cd, by = 'organism_part')
```


Looking into the heatmaps and PCA plots, what is the driving factor of the clustering of samples?

Do the samples cluster by the donor (individual)? or do they cluster by the type of tissue (normal, tumor, metastatic?) 


# Stylizing your HTML report

Check out [this page](https://bookdown.org/yihui/rmarkdown/html-document.html) that describe how to stylize your rmarkdown reports. For example try adding the following lines at the top of the page: 

Replace this line

```
output: html_document
```

with this:
```
output:
  html_document:
    theme: united
    highlight: tango
```

Re-compile your report and see how it looks. 

- How can you also keep your code but keep it folded by default in the output?
- How can you have a floating table of contents on your screen? 



